---
title: "EDA-challenge.qmd"
format: html
editor: visual
---

### Body Size dimorphism (BSD)

Creating a new variable which is the ratio of average male to female body mass

```{r}
d$BSD <- d$Body_mass_male_mean/d$Body_mass_female_mean
d$BSD
```

### Sex_Ratio

creating a new variable sex ratio, which is the ratio of the number of adult females to adult males in a typical group.

```{r}
d$sex_ratio<-d$AdultFemale/d$AdultMales
d$sex_ratio
```

### Defensibility Index DI

ratio of day range length to the diameter of the home range

```{r}
radius<-sqrt(d$HomeRange_km2/pi)
diameter<-2*radius
d$DI<-d$DayLength_km/diameter
d$DI
```

### 4. Plotting the relationship between day range length (y- axis) and time spent moving (x-axis)

#### Family

```{r}
library(skimr)
library(kableExtra)
f<-"https://raw.githubusercontent.com/difiore/ada-datasets/main/data-wrangling.csv"
d <- read.table(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
p <- ggplot(data = d, aes(x = log(Move), y = log(DayLength_km), 
                          color = factor(Family)))
p <- p + xlab("log(Time Spent Moving)") + ylab("log(Day Range Length)")
p <- p + geom_point(na.rm = TRUE)
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p <- p + facet_wrap(~Family, ncol = 4)
p <- p + theme(legend.position = "none")
p
```

```{r}
ggplot(d, aes(x= Move, y = DayLength_km))+
  geom_point()+
  geom_smooth(mmethod = "lm", se = FALSE, color = "black")+
  labs(x="Time spent Moving", y = "Day Range Length")+
  theme_minimal()+
  facet_wrap(~Family, scales = "free")
```

#### Overall

```{r}
library(skimr)
library(kableExtra)
f<-"https://raw.githubusercontent.com/difiore/ada-datasets/main/data-wrangling.csv"
```

```{r}
ggplot(d, aes(x= Move, y = DayLength_km))+
  geom_point()+
  geom_smooth(mmethod = "lm", se = FALSE, color = "black")+
  labs(x="Time spent Moving", y = "Day Range Length")+
  theme_minimal()
```

##### questions

1.) Species that spend more time mvoing do not travel farther overall 2.) withing a particular primate family cebidae seem to move alot but don't travel a great distance

### 5. Plot the Relationship between Day Range Length (y axis) and group size (x axis)

#### Overall

```{r}
ggplot(d, aes(x = MeanGroupSize, y = DayLength_km))+
geom_point()+
geom_smooth(method = "lm", se = FALSE, color = "blue")+
labs(x = "Group Size", y = "Day Range Length (km)")+
 theme_minimal()
```

#### Family

```{r}
ggplot(d, aes(x = MeanGroupSize, y = DayLength_km))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "pink")+
  labs(x = "Group Size", y = "Day Range Length (km)")+
  theme_minimal()+
  facet_wrap(~Family, scales = "free")
```

###### Questions

1.) Species that live in larger groups do travel father overall. 2.) looks like the hominidae trravel far the greater the group size. \### 6. Plot the relationship between canine size dimorphism and body size dimorphism

#### family

```{r}
d$BSD <- d$Body_mass_male_mean / d$Body_mass_female_mean
d <- d[!is.infinite(d$BSD) & !is.na(d$BSD),]

ggplot(d, aes(x = BSD, y = Canine_Dimorphism))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "purple")+
  labs(x= "Body Size Dimorphism", y = "Canine Size Dimorphism")+
  theme_minimal()+
  facet_wrap(~Family, scales = "free")
```

#### Overall

```{r}
d$BSD <- d$Body_mass_male_mean / d$Body_mass_female_mean
d <- d[!is.infinite(d$BSD) & !is.na(d$BSD),]

ggplot(d, aes(x = BSD, y = Canine_Dimorphism))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "purple")+
  labs(x= "Body Size Dimorphism", y = "Canine Size Dimorphism")+
  theme_minimal()
```

##### Questions

1.) it does look like the greater the body size the greater the canine size (I do not understand "transform these variables"

### 7. Diet_Strategy Variable

```{r}
library(dplyr)
d <- d %>%
  mutate(diet_strategy = case_when (
    Fruit >= 50 ~ "frugivore",
    Leaves >=50 ~ "folivore",
    Fruit < 50 & Leaves < 50 ~"omnivore",
    TRUE ~ NA_character_
  )) 
ggplot(d %>% filter(!is.na(diet_strategy)), aes(x=diet_strategy, y = MeanGroupSize))+
  geom_boxplot()+ 
  labs(x="Diet Strategy", y = "Group Size")+ 
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

##### Questions

1.) Yes Fugiovre do live in bigger size than folivore 

### 8.

library(dplyr)

```{r}
d %>%
  mutate(Binomial = paste(Genus, Species)) %>%
  select(Binomial, Family, Brain_Size_Species_Mean, Body_mass_male_mean) %>%
  summarise(
    avg_brain_size = mean(Brain_Size_Species_Mean, na.rm = TRUE),
    avg_body_mass = mean(Body_mass_male_mean, na.rm = TRUE) 
  ) %>%
  arrange(avg_brain_size) %>%
  print()
```
